---
title: "uber analysis"
author: "Nikhil"
date: "2023-02-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installing & calling the packages 
```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggthemes)
library(tidyr)
library(scales)
library(DT)
```

# load the data and create your first join :)
```{r}
apr_data <- read.csv("uber-raw-data-apr14.csv")
may_data <- read.csv("uber-raw-data-may14.csv")
jun_data <- read.csv("uber-raw-data-jun14.csv")
jul_data <- read.csv("uber-raw-data-jul14.csv")
aug_data <- read.csv("uber-raw-data-aug14.csv")
sep_data <- read.csv("uber-raw-data-sep14.csv")

data_2014 <- rbind(apr_data, may_data, jun_data, jul_data, aug_data, sep_data)
```

# Now we Clean the data 
1. Posixct encodes the date and time as number of seconds from 1970
2. Posixlt stores date and time as a list of month, date and time 
3. Bote the Posix formats allow us subtraction between 2 dates 
```{r}
#Make data set familiar with format using POSIXct
data_2014$Date.Time <- as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S")

#Done to divide time column 
data_2014$Time <- format(as.POSIXct(data_2014$Date.Time, format = "%m/%d/%y %H:%M:%S"), format = "%H:%M:%S")
data_2014$Date.Time <- ymd_hms(data_2014$Date.Time)

# Now we divide by month, day, year 
data_2014$Day <- factor(day(data_2014$Date.Time))
data_2014$Month <- factor(month(data_2014$Date.Time))
data_2014$year <- factor(year(data_2014$Date.Time))
data_2014$Day_of_week <- factor(wday(data_2014$Date.Time, label = TRUE))

#Now we divide the time 
data_2014$hour <- factor(hour(data_2014$Date.Time))
data_2014$minute <- factor(minute(data_2014$Date.Time))
data_2014$second <- factor(second(data_2014$Date.Time))
```



# Plotting data for number of trips in an hour 
```{r, message=FALSE}
hour_table2 <- data_2014 %>% 
  group_by(hour) %>% 
  dplyr::summarize(total = n())

ggplot(hour_table2, aes(hour, total)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "red") +
  ggtitle("Maximum number of trips taken between 5:00 - 6:00 Pm")+
  theme(legend.position = "none")+
  scale_y_continuous(labels = comma)
 
```



# Plotting trips by hour and month 
```{r, message=FALSE}
month_tab <- data_2014 %>% 
  group_by(Month, hour) %>% 
  summarize(total = n())

ggplot(month_tab, aes(hour, total, fill = Month)) +
  geom_bar(stat = "identity", ) +
  ggtitle("trips by month") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = comma) 
```


# Plotting by trips taken every day 
```{r, message=FALSE}
day_tab <- data_2014 %>% 
  group_by(Day) %>% 
  summarize(total = n())

ggplot(day_tab, aes(Day, total)) +
  geom_bar(stat = "identity", fill = "steelblue" ,color = "red")+
  ggtitle("trips taken every day")+
  theme(legend.position = "none")
```


# Plotting trips every day of the month (Dodge bar chart)
```{r, message=FALSE}
day_month_tab <- data_2014 %>% 
  group_by(Month, Day_of_week) %>% 
  summarize(total = n())

ggplot(day_month_tab, aes(Month, total, fill = Day_of_week))+
  geom_bar(stat = "identity", position = "dodge")
```

# Number of Trips taking place during months in a year
```{r, message=FALSE}
month_year_tab <- data_2014 %>% 
  group_by(year, Month) %>% 
  summarize(total = n())

ggplot(month_year_tab, aes(Month, total)) + 
  geom_bar(stat = "identity", fill = "steelblue", color = "red") +
  ggtitle("Number of Trips taking place during months in a year") +
  theme(legend.position = "none") +
  scale_fill_manual(values = colors)
```

# trips of month by weekday 
```{r, message=FALSE}
month_weekday <- data_2014 %>% 
  group_by(Month, Day_of_week) %>% 
  summarize(total = n())

ggplot(month_weekday, aes(Month, total, fill = Day_of_week)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("trips of month by weekday") +
  theme(legend.position = "none")

```

# heatmap representing rides taken on different day of week in a month
```{r, message=FALSE}
ggplot(month_weekday, aes(Month, Day_of_week, fill = total)) +
  geom_tile(color ="white") +
  ggtitle("heatmap representing rides taken on different day of week in a month")

```



